---
title: "Stat. 652: Final_pe7112"
author: "Prof. Eric A. Suess"
format: 
  html:
    self-contained: true
---

Answer two questions:

1.  What are the important variables identified by the Boruta algorithm from the **Ozone** data?

**Answer:**

**Important attributes:** v1,v4,v5,v7,v8,v9,v10,v11,v12,v13.

**Unimportant attributes:** v2 , v3 , v6.

Boruta.Short got 2 tentative attributes v2,v6 due to less number of iterations. After applying TentativeRoughFix function these are classified as unimportant attributes

2.  What are the important variables identified by the Boruta algorithm from the **titanic training** data?

**Answer:**

**Important attributes:** "pclass", "sex" "age" "sib_sp" "parch" "ticket" "fare" "cabin" "embarked".

**Unimportant attributes:** "PassengerId".

**Tentative attributes:** "Name"

After applying TentativeRoughFix function the "name" attribute is classified under important attributes.

```{r}
library(pacman)
p_load(tidyverse, janitor, naniar, DataExplorer, Boruta, mlbench)
```

```{r}
data(Ozone)

head(Ozone)
```

Note that the target variable is V4 = Daily maximum one-hour-average ozone reading.

Convert V1, V2, and V3 to integers.

```{r}
Ozone <- Ozone %>% mutate(
  V1 = as.integer(V1),
  V2 = as.integer(V2),
  V3 = as.integer(V3)
)

head(Ozone)
```

```{r}
Ozone2 <- Ozone %>% clean_names()

head(Ozone2)
```

It is always a good idea to check for duplicate records/examples/rows in your dataset.

```{r}
get_dupes(Ozone2)
```

Start by investigating the missing values and completeness of the features in the data. Note that the *age* variable contains some missing values.

```{r}
vis_miss(Ozone2)
gg_miss_var(Ozone2)
gg_miss_var(Ozone2, show_pct = TRUE)
```

Drop V9 because it is missing more than 40% of its values.

```{r}
Ozone2 <- Ozone2 %>% select(-v9)

vis_miss(Ozone2)
gg_miss_var(Ozone2)
gg_miss_var(Ozone2, show_pct = TRUE)
```

```{r}
Ozone2 <- drop_na(Ozone2)

vis_miss(Ozone2)
gg_miss_var(Ozone2)
gg_miss_var(Ozone2, show_pct = TRUE)
```

```{r}
Boruta.Ozone <- Boruta(v4 ~ ., data = Ozone2, doTrace = 2, ntree = 500)
```

```{r}
Boruta.Ozone
```

```{r}
plot(Boruta.Ozone)
```

```{r}
Boruta.Short <- Boruta(v4 ~ ., data = Ozone2, maxRuns = 12)
```

```{r}
Boruta.Short 
```

```{r}
plot(Boruta.Short)
```

```{r}
TentativeRoughFix(Boruta.Short)
```

```{r}
getConfirmedFormula(Boruta.Ozone)
```

```{r}
attStats(Boruta.Ozone)
```

```{r}
plotImpHistory(Boruta.Ozone)
```

```{r}
library(titanic)

data(titanic_train)

```

```{r}

library(tidyverse) 
library(janitor)   
library(naniar)    
library(Boruta)    
```

```{r}

titanic_train <- titanic_train %>%
  mutate(
    Survived = as.factor(Survived),
    Pclass = as.factor(Pclass),
    Sex = as.factor(Sex),
    Embarked = as.factor(Embarked),
    Age = as.numeric(Age), 
    SibSp = as.integer(SibSp),
    Parch = as.integer(Parch),
    Fare = as.numeric(Fare)
  ) %>%
  clean_names() 

```

```{r}
# Check for duplicates
if (any(duplicated(titanic_train))) {
  titanic_train <- titanic_train[!duplicated(titanic_train), ]
}
```

```{r}
# Visualize missing values
vis_miss(titanic_train)
```

```{r}
titanic_train <- drop_na(titanic_train)
```

```{r}
set.seed(123) # Set seed 
boruta_output <- Boruta(survived ~ ., data = titanic_train, doTrace = 0)
```

```{r}
print(boruta_output)
```

```{r}
plot(boruta_output)
```

```{r}
TentativeRoughFix(boruta_output)
```

```{r}
boruta_stats <- attStats(boruta_output)
confirmed_vars <- rownames(boruta_stats[boruta_stats$decision == 'Confirmed', ])
print(confirmed_vars)
```

```{r}
attStats(boruta_output)
```

```{r}
plotImpHistory(boruta_output)
```
